Refinement for open automata

### 2021-05-04 2021-Refinement/report.tex
some proofreading and comments

Throughout this paper, tuples might be noted differently depending on what they represent.
Quantification will be used loosely, any bound name in a quantification means that the object is constrained to have that previous value, and any new name is bound by the quantifier.
For instance \(\forall \mpar{x, y}, \exists \mpar{x', y}, P\) means \(\forall \mpar{x, y}, \exists \mpar{x', y'}, y = y' \wedge P\).

Family of values, or equivalently maps will be noted \(\mset{i \mapsto x_i}{i \in I}\), \(\mset{i \gets x_i}{i \in I}\) or \(x_i^{i \in I}\).
The latter will only be used when unambiguous; for instance \(\mpar{ax}^{x \in \setR}\) represents a scaling function, \(c^{i \in I}\) is a constant function over \(I\), but \(\mbrc{\alpha \mapsto 1, \beta \mapsto 2, \gamma \mapsto 3}\) must be represented with one of the two first notation.
The disjoint union of two maps \(\varphi: I \to X\) and \(\psi: J \to Y\) with \(I \cap J = \emptyset\) is \(\varphi \uplus \psi: I \uplus J \to X \cup Y\).

For a relation \(\wrel{}{}{}\) which uses a property \(P\) to prove that two element are related \(\forall x \, x', x \wrel{}{}{} x' \iff \exists w: P\mpar{x, x'}\), the fact that there is a valid witness \(w: P\mpar{x, x'}\) is noted \(\wrel{x}{x'}{w}\).

@@

The terms \(\terms\) is a term algebra.
The action algebra \(\actions\) is the set of pairs \(\labels \times \terms\) where \(\labels\) are the action labels.
The formulae \(\formulae\) are at least the first order formulae over \(\terms\) and \(\labels\).
\end{defi}
The term algebra and action labels are arbitrary.
An element \(\mpar{l, t} \in \actions\) is noted \(l\mpar{t}\).
The formulae form at least a first order logic with an equality relation that is not necessarily the equality on \(\terms\).

@@

\begin{defi}[Interpretation, Values, Satisfiability, (Parallel) substitution]
We assume that the following are given:
\defitem The values \(\values\), which are interpretations of closed terms.
\defitem The satisfiability relation on closed formulae, \({\vdash} \subseteq \rformulae\).
\defitem The substitution in \(e \in E\) of \(x \in \fvars{e}\) by \(t \in \terms\), \(e\subst{t}{x}\).
\defitem The parallel substitution in \(e \in E\) of variables in \(V\) by \(\psi: V \to \terms\), \(e\psubst{\psi}\).
\end{defi}
For the parallel substitution, the set \(V\) is not required to be a subset of \(\fvars{e}\).
In the case where it is not, the variables in \(V \setminus \fvars{e}\) are not substituted.
The substitutions might give a ill-formed expression; for instance let the terms be integers and pairs with (pointwise) addition, \(\mpar{a + b}\psubst{a \mapsto 7, b \mapsto \mpar{4, 5}}\) is a ill-formed term.
This can be guarded with the check \(e\subst{t}{x} \in E\) and \(e\psubst{\psi} \in E\) and it will implicitely be the case, for instance when there is quantification on \(t\) and \(\psi\), to simplify notations.

@@

@@

To illustrate these definitions let's look at two implementations (figure \ref{fig:enable}) of the LOTOS\marginpar{cite something here} operator enable in the open automata model.
The enable operator runs its the left hand side agent until it chooses to finish, at which point it produces an action \(\delta\) that is synchronised with the first action of the right hand side agent which must be \(\act{acc}\), then only the latter agent runs.
During the synchronised action \Quentin{data can be exchanged}{I need to read a little bit more on that; is that a return code?}.
\begin{exi}[Enable, state-oriented]
\begin{figure}
\centering
\input{enable_state.tex}
\vrule
\input{enable_var.tex}
\caption{Enable operator implementation with open automata, on the left state oriented, on the right variable oriented}
\label{fig:enable}
\end{figure}
To draw automata, the usual circle for states and simple arrows for transitions convention is used.
The initial state is indicated by a double circle.
States names are indicated inside the circles and transitions labels are drawn near their corresponding arrow.

The open transitions do not indicate the source and target states since that is the role of the transitions arrows; only the action is on the bottom side of the open transition.
Also in the example, transitions are specified using variables in place of actions.
This is used to represent all the transitions created by a correct substitution of these variables, which cannot be drawn otherwise as there are often infinitly many.

Let's get back to the example.
The automata on the left side is \(\OA{\mbrc{L, R}}{L}{\emptyset}{\mbrc{}}{\mbrc{l, r}}{T}\) where transitions in \(T\) are generated by
\begin{align*}
	\OT{L}{L}{x}{\mbrc{l \mapsto x}}{x \neq \delta}{\mbrc{}} &&
	\OT{L}{R}{\tau}{\mbrc{l \mapsto \delta\mpar{x}, r \mapsto \act{acc}\mpar{y}}}{x = y}{\mbrc{}} &&
	\OT{R}{R}{x}{\mbrc{r \mapsto x}}{\top}{\mbrc{}}
\end{align*}
Note that we could have expressed the transition in the middle with \(\mbrc{l \mapsto \delta\mpar{x}, r \mapsto \act{acc}\mpar{x}}\) which would have avoided many transitions with trivially false guards like the one where \(x \mapsto 1, y \mapsto 2\) which has the guard \(1 = 2\).

This automaton is an implementation of the enable operator because it begins in the state \(L\) where it allows any transition from its hole \(l\) which is not \(\delta\) then synchronises its holes on the same data, effectively allowing a data exchange when it goes into state \(R\), and finally allows any transition from its hole \(r\).
Only the synchronisation is hidden to the exterior by sending a \(\tau\) which is usually a silent action always allowed, unsynchronised with other holes and passed unmodified.
Well here \(\tau\) is not always allowed from the holes in the state \(L\) the hole \(r\) cannot emmit it for the sake of simplifying the first example of open automaton.
\end{exi}
\begin{exi}[Enable, variable-oriented]
The automata drawn on the right side of the figure is an alternative implementation of the enable operator which is FH-bisimilar.
The drawing introduces the double arrows which is used to denote the initial valuation.
The automata drawn is \(\OA{\mbrc{\_}}{\_}{\mbrc{v}}{l^\mbrc{v}}{\mbrc{l, r}}{T}\) where transitions in \(T\) are like previously all the transitions generated by
\begin{align*}
	\OT{\_}{\_}{x}{\mbrc{l \mapsto x}}{v = l \wedge x \neq \delta}{\mbrc{}} &&
	\OT{\_}{\_}{\tau}{\mbrc{l \mapsto \delta\mpar{x}, r \mapsto \act{acc}\mpar{y}}}{v = l \wedge x = y}{\mbrc{v \gets r}} &&
	\OT{\_}{\_}{x}{\mbrc{r \mapsto x}}{v = r}{\mbrc{}}
\end{align*}
Note that in this example \(r\) is a hole name and also a closed term, but not a variable.
The same applies for \(l\).
\(v\) is a variable name for the automaton and a variable in the guard, so its value is not substituted.
For instance \nmm{\OT{\_}{\_}{\tau}{\mbrc{l \mapsto \tau}}{v = l \wedge \tau \neq \delta}{\mbrc{}}} is in \(T\) but not \nmm{\OT{\_}{\_}{\tau}{\mbrc{l \mapsto \tau}}{l = l \wedge \tau \neq \delta}{\mbrc{}}} (result of the substitution \(v \mapsto l\)).

@@

\section{Semantic and composition of Open Automata}\label{sec:semcompo}
The interpretation of open automata is a partially specified automaton, where the partially specification comming from the holes.
This specification can be completed by composing automata into holes so it is natural to take a look at the composition process.
This was already defined on pNets in previous work \cite{henrio:01299562} but never completely formalised on open automata.
The below definition of composition is a direct translation of composition on pNets for open automata.
\begin{defi}[Composition of open automata]
The composition of \(A_c \defobject \OA{S_c}{s_{0c}}{V_c}{\sigma_{0c}}{J_c}{T_c}\) in the hole \(k \in J_p\) of \(A_p \defobject \OA{S_p}{s_{0p}}{V_p}{\sigma_{0p}}{J_p}{T_p}\) is
\begin{align*}
	A_p\subst{A_c}{k} \defnotation & \OA{S_p \times S_c}{\mpar{s_{0p}, s_{0c}}}{V_p \uplus V_c}{\sigma_{0p} \uplus \sigma_{0c}}{J_c \uplus J_p \setminus \mbrc{k}}{T} \\
	\text{With } T \defobject & \mset{\OT{\mpar{s_p, s_c}}{\mpar{s'_p, s'_c}}{\alpha_p}{\beta_j^{j \in J'_c \uplus J'_p \setminus \mbrc{k}}}{g_p \wedge g_c \wedge \alpha_c = \beta_k}{\psi_p \uplus \psi_c}}{\OTx{p}{}{}{p} \in T_p, \OTx{c}{}{}{c} \in T_c} \\
	& \cup \mset{\OT{\mpar{s_p, s_c}}{\mpar{s'_p, s_c}}{\alpha_p}{\beta_j^{j \in J'_p}}{g_p}{\psi_p}}{\OTx{p}{}{}{p}, k \notin J', s_c \in S_c}
\end{align*}
Similarly to expression substitution, the parallel composition is noted \(A\psubst{A_j^{j \in J}}\).
\end{defi}
The actions emitted when \(A_c\) makes a transition is sychronised with the action of the hole \(k\) in transitions of \(A_p\) which have it as a hole actions (first transition set, \(\alpha_c = \beta_k\)).
No transition of \(A_c\) is performed when a transition that do not refer to the hole \(k\) is performed in \(A_p\) (second transition set, \(k \notin J'\)).
The composition may look like a handshake, with both automata running in parallel (product of states and joint variables) but it is actually not symmetric.
\begin{exi}
This example is derived and adapted from a traffic light controller in a collection of examples for pNets.
It is supposed to be a single traffic light.

The figure \ref{fig:tls} is the light controller with some synchronisation logic.
It still takes an unimplemented control circuit in the hole \(C\) which gives the timings and an also unimplemented register in the hole \(R\) to count external \(\act{tick}\) actions.
The three states are used to remember which colored light is on and this color can be retrieved by the environment by synchronising with the actions \(\act{onXxx}\) if it is not stored externally.

@@

\begin{figure}
\centering
\input{Traffic_Lights_Spec.tex}
\caption{The specification of a traffic light system}
\label{fig:tls}
\end{figure}
\begin{figure}
\centering
\input{Traffic_Lights_Controller.tex}
\vrule
\input{Traffic_Lights_Register.tex}
\caption{On the left: The controller agent; On the right: The counter agent}
\label{fig:tlh}
\end{figure}
The component we choose to compose in the holes are in figure \ref{fig:tlh}.
On the left is the controller agent which will be composed in the hole \(C\).
Its role is to decide the duration before switching the lights.
This implementation is simply a constant time for each color.

@@

The automaton on figure \ref{fig:tlf} is bisimilar to the composition of the specification on figure \ref{fig:tls} and the agents on figure \ref{fig:tlh}.
The result of the composition has 36 states (\(3 \times 2 \times 6\)) but only 6 are reachable from the initial configuration.
Also satisfiable guards were replaced with simplified equivalent.

The transitions obtained by composition are the following (generic and simplified version):
\begin{itemize}
\item \nmm{\OT{XS}{XC}{\tau}{\mbrc{}}{\top}{\mbrc{t \gets k, c \gets 0}}} which is the composition of \nmm{\OT{X}{X}{\tau}{\mbrc{C \mapsto \theta\mpar{x}, R \mapsto \act{set}\mpar{x}}}{\top}{\mbrc{}}} from the specification with \nmm{\OT{S}{C}{\act{set}\mpar{x}}{\mbrc{}}{\top}{\mbrc{t \gets x, c \gets 0}}} from the register and \nmm{\OT{n}{n+1}{\theta\mpar{k}}{\mbrc{}}{\top}{\mbrc{}}} from the controller.
\item \nmm{\OT{XC}{XC}{\act{tick}}{\mbrc{}}{c < t}{\mbrc{c \gets c + 1}}} which is the composition of \nmm{\OT{X}{X}{\act{tick}}{\mbrc{R \mapsto \act{tick}}}{\top}{\mbrc{}}} from the specification with \nmm{\OT{C}{C}{\act{tick}}{\mbrc{}}{c < t}{\mbrc{c \gets c + 1}}} from the register.
\item \nmm{\OT{XC}{ZS}{\act{TurnXxx}}{\mbrc{}}{c = t}{\mbrc{}}} which is the composition of \nmm{\OT{X}{Z}{\act{TurnXxx}}{\mbrc{R \mapsto \act{over}\mpar{x}, C \mapsto \delta\mpar{x}}}{\top}{\mbrc{}}} from the specification with \nmm{\OT{C}{S}{\act{over}\mpar{c}}{\mbrc{}}{c = t}{\mbrc{}}} from the register and \nmm{\OT{X}{Z}{\delta\mpar{x}}{\mbrc{}}{\top}{\mbrc{}}} from the controller.
\end{itemize}

>>>>

Throughout this paper, tuples might be noted differently depending on what they represent.
Quantification will be used loosely, any bound name in a quantification means that the object is constrained to have that previous value, and any new name is bound by the quantifier.
For instance \(\forall \mpar{x, y}, \exists \mpar{x', y}, P\) means \(\forall \mpar{x, y}, \exists \mpar{x', y'}, y = y' \wedge P\).
\Ludo{Strange}{do we really need this interpretation of quantifiers? I find it disturbing}

Family of values, or equivalently maps will be noted \(\mset{i \mapsto x_i}{i \in I}\), \(\mset{i \gets x_i}{i \in I}\) or \(x_i^{i \in I}\).
The latter will only be used when unambiguous; for instance \(\mpar{ax}^{x \in \setR}\) represents a scaling function, \(c^{i \in I}\) is a constant function over \(I\), but \(\mbrc{\alpha \mapsto 1, \beta \mapsto 2, \gamma \mapsto 3}\) must be represented with one of the two first notation.
\Ludo{Strange}{ you mean the last notation is forbidden? but isn't it introduced before?}
The disjoint union of two maps \(\varphi: I \to X\) and \(\psi: J \to Y\) with \(I \cap J = \emptyset\) is \(\varphi \uplus \psi: I \uplus J \to X \cup Y\).

For a relation \(\wrel{}{}{}\) which uses a property \(P\) to prove that two element are related \(\forall x \, x', x \wrel{}{}{} x' \iff \exists w: P\mpar{x, x'}\), the fact that there is a valid witness \(w: P\mpar{x, x'}\) is noted \(\wrel{x}{x'}{w}\).
\Ludo{Strange}{not clear: it seems that $w$ is useless  I guess you mean $P$ depend on $w$ but it is not exactly what is written}

@@

The terms \(\terms\) is a term algebra.
The action algebra \(\actions\) is the set of pairs \(\labels \times \terms\) where \(\labels\) are the action labels.
The formulae \(\formulae\) are at least the first order formulae over \(\terms\) and \(\labels\).
\end{defi}
The term algebra and action labels are arbitrary.
An element \(\mpar{l, t} \in \actions\) is noted \(l\mpar{t}\).
The formulae contain at least first order logic terms with an equality relation that is not necessarily \Ludo{?}{restricted to?} the equality on \(\terms\).

@@

\begin{defi}[Interpretation, Values, Satisfiability, (Parallel) substitution]
We assume that the following are given:
\defitem The values \(\values\), which are interpretations of closed terms.
\defitem The satisfiability relation on closed formulae, \({\vdash} \subseteq \rformulae\).
\defitem The substitution in \(e \in E\) of \(x \in \fvars{e}\) by \(t \in \terms\), \(e\subst{t}{x}\).
\defitem The parallel substitution in \(e \in E\) of variables in \(V\) by \(\psi: V \to \terms\), \(e\psubst{\psi}\).
\end{defi}
For the parallel substitution, the set \(V\) is not required to be a subset of \(\fvars{e}\).
In the case where it is not, the variables in \(V \setminus \fvars{e}\) are not substituted.
The substitutions might give a ill-formed expression; for instance let the terms be integers and pairs with (pointwise) addition, \(\mpar{a + b}\psubst{a \mapsto 7, b \mapsto \mpar{4, 5}}\) is a ill-formed term.
This can be guarded with the check \(e\subst{t}{x} \in E\) and \(e\psubst{\psi} \in E\) and it will implicitly be the case, for instance when there is quantification on \(t\) and \(\psi\), to simplify notations.

@@

\Ludo{?}{I do not understand the meaning of the net paragraph: I must have missed something}

@@

To illustrate these definitions we consider  two implementations (figure \ref{fig:enable}) of the LOTOS\marginpar{cite something here} operator enable in the open automata model.
The enable operator runs  the left hand side agent until it chooses to finish, at which point it produces an action \(\delta\) that is synchronised with the first action of the right hand side agent which must be \(\act{acc}\), then only the latter agent runs.
During the synchronised action \Quentin{data can be exchanged}{I need to read a little bit more on that; is that a return code?}.

\begin{exi}[Enable, state-oriented]
\begin{figure}
\centering
\input{enable_state.tex}
\vrule
\input{enable_var.tex}
\caption{Enable operator implementation with open automata, on the left state oriented, on the right variable oriented}
\label{fig:enable}
\end{figure}

Graphical notations for showing automata are as follows.
Automata are shown in the figure, the usual circle for states and simple arrows for transitions convention is used.
The initial state is indicated by a double circle.
States names are indicated inside the circles and transitions labels are drawn near their corresponding arrow.
The open transitions do not indicate the source and target states since that is the role of the transitions arrows; only the action is on the bottom side of the open transition.
Also in the example, transitions are specified using variables in place of actions.
This is used to represent all the transitions created by a correct substitution of these variables, which cannot be drawn as there are often infinitely many.

We can now explain the enable operator example.
The automata on the left side is \(\OA{\mbrc{L, R}}{L}{\emptyset}{\mbrc{}}{\mbrc{l, r}}{T}\) where transitions in \(T\) are generated by
\begin{align*}
	\OT{L}{L}{x}{\mbrc{l \mapsto x}}{x \neq \delta}{\mbrc{}} &&
	\OT{L}{R}{\tau}{\mbrc{l \mapsto \delta\mpar{x}, r \mapsto \act{acc}\mpar{y}}}{x = y}{\mbrc{}} &&
	\OT{R}{R}{x}{\mbrc{r \mapsto x}}{\top}{\mbrc{}}
\end{align*}
Note that we could have expressed the transition in the middle with \(\mbrc{l \mapsto \delta\mpar{x}, r \mapsto \act{acc}\mpar{x}}\) which would have avoided many transitions with trivially false guards like the one where \(x \mapsto 1, y \mapsto 2\) which has the guard \(1 = 2\).

This automaton is an implementation of the enable operator because it begins in the state \(L\) where it allows any transition from its hole \(l\) which is not \(\delta\) then the automaton synchronises its holes on the same data, effectively allowing a data exchange when it goes into state \(R\), and finally allows any transition from its hole \(r\).
We use the standard convention that uses $\tau$ as a non-observable transition, never synchronised with other actions and passed unmodified.
Here the synchronisation is hidden to the exterior by sending a \(\tau\).
\Ludo{Well}{avoid informal English in scientific writing} here \(\tau\) is not always allowed from the holes in the state \(L\) the hole \(r\) cannot emit it for the sake of simplifying the first example of open automaton.
\end{exi}
\begin{exi}[Enable, variable-oriented]
The automata drawn on the right side of the figure is an alternative implementation of the enable operator which is FH-bisimilar to the previous one\footnote{FH-bisimulation is defined in ~\cite{addcitation}}.
The drawing introduces the double arrows which is used to denote the initial valuation.
The automata drawn is \(\OA{\mbrc{\_}}{\_}{\mbrc{v}}{l^\mbrc{v}}{\mbrc{l, r}}{T}\) where transitions in \(T\) are like previously all the transitions generated by
\begin{align*}
	\OT{\_}{\_}{x}{\mbrc{l \mapsto x}}{v = l \wedge x \neq \delta}{\mbrc{}} &&
	\OT{\_}{\_}{\tau}{\mbrc{l \mapsto \delta\mpar{x}, r \mapsto \act{acc}\mpar{y}}}{v = l \wedge x = y}{\mbrc{v \gets r}} &&
	\OT{\_}{\_}{x}{\mbrc{r \mapsto x}}{v = r}{\mbrc{}}
\end{align*}
Note that in this example \(r\) is a hole name and also a closed term, but not a variable.
The same applies for \(l\).
\(v\) is a variable name for the automaton and a variable in the guard, \Ludo{so its value is not substituted}{I do not understand this: it seems to mean that the substitution of v has no effect}.\Ludo{?}{I believe this needs to be explained more clearly, perhaps based on the def 6 below}
For instance \nmm{\OT{\_}{\_}{\tau}{\mbrc{l \mapsto \tau}}{v = l \wedge \tau \neq \delta}{\mbrc{}}} is in \(T\) but not \nmm{\OT{\_}{\_}{\tau}{\mbrc{l \mapsto \tau}}{l = l \wedge \tau \neq \delta}{\mbrc{}}} (result of the substitution \(v \mapsto l\)).

@@

\section{Semantic and composition of Open Automata}\label{sec:semcompo}
The interpretation of open automata is a partially specified automaton, where the partial specification comes from the holes.
This specification can be completed by composing automata into holes so it is natural to take a look at the composition process.
This was already defined on pNets in previous work \cite{henrio:01299562} but never completely formalised on open automata.
The  definition of composition below is a direct translation of composition on pNets for open automata.
\begin{defi}[Composition of open automata]
The composition of \(A_c \defobject \OA{S_c}{s_{0c}}{V_c}{\sigma_{0c}}{J_c}{T_c}\) in the hole \(k \in J_p\) of \(A_p \defobject \OA{S_p}{s_{0p}}{V_p}{\sigma_{0p}}{J_p}{T_p}\) is
\begin{align*}
	A_p\subst{A_c}{k} \defnotation & \OA{S_p \times S_c}{\mpar{s_{0p}, s_{0c}}}{V_p \uplus V_c}{\sigma_{0p} \uplus \sigma_{0c}}{J_c \uplus J_p \setminus \mbrc{k}}{T} \\
	\text{With } T \defobject & \mset{\OT{\mpar{s_p, s_c}}{\mpar{s'_p, s'_c}}{\alpha_p}{\beta_j^{j \in J'_c \uplus J'_p \setminus \mbrc{k}}}{g_p \wedge g_c \wedge \alpha_c = \beta_k}{\psi_p \uplus \psi_c}}{\OTx{p}{}{}{p} \in T_p, \OTx{c}{}{}{c} \in T_c} \\
	& \cup \mset{\OT{\mpar{s_p, s_c}}{\mpar{s'_p, s_c}}{\alpha_p}{\beta_j^{j \in J'_p}}{g_p}{\psi_p}}{\OTx{p}{}{}{p}, k \notin J', s_c \in S_c}
\end{align*}
Similarly to expression substitution, the parallel composition is noted \(A\psubst{A_j^{j \in J}}\).
\end{defi}
The actions emitted when \(A_c\) makes a transition is sychronised with the action of the hole \(k\) in transitions of \(A_p\) which have it as a hole actions (first transition set, \(\alpha_c = \beta_k\)).
No transition of \(A_c\) is performed when a transition that do not refer to the hole \(k\) is performed in \(A_p\) (second transition set, \(k \notin J'\)).
The composition may look like a handshake, with both automata running in parallel (product of states and joint variables) but it is actually not symmetric.
\begin{exi}
This example is derived and adapted from a traffic light controller in a collection of examples for pNets.
It is supposed to be a single traffic light.

\Ludo{Figure \ref{fig:tls} shows}{rephrased: typo} the light controller with some synchronisation logic.
It  takes an unimplemented control circuit in the hole \(C\) which gives the timings and an also unimplemented register in the hole \(R\) to count external \(\act{tick}\) actions.
The three states are used to remember which colored light is on and this color can be retrieved by the environment by synchronising with the actions \(\act{onXxx}\) if it is not stored externally.

@@

\begin{figure}
\centering
\input{Traffic_Lights_Spec.tex}
\caption{The specification of a traffic light system}
\label{fig:tls}
\end{figure}
\begin{figure}
\centering
\input{Traffic_Lights_Controller.tex}
\vrule
\input{Traffic_Lights_Register.tex}
\caption{On the left: The controller agent; On the right: The counter agent}
\label{fig:tlh}
\end{figure}
The component we choose to compose in the holes are in \Ludo{F}{}igure \ref{fig:tlh}.
On the left is the controller agent which will be composed in the hole \(C\).
Its role is to decide the duration before switching the lights.
This implementation is simply a constant time for each color.

@@

The automaton on Figure \ref{fig:tlf} is bisimilar to the composition of the specification on Figure \ref{fig:tls} and the agents on Figure \ref{fig:tlh}.
The result of the composition has 36 states (\(3 \times 2 \times 6\)) but only 6 are reachable from the initial configuration.
The transitions obtained by composition are the following (generic and simplified version). \Ludo{Also satisfiable guards were replaced with simplified equivalent}{moved but a rephrasing may be better}:
\begin{itemize}
\item \nmm{\OT{XS}{XC}{\tau}{\mbrc{}}{\top}{\mbrc{t \gets k, c \gets 0}}} which is the composition of \nmm{\OT{X}{X}{\tau}{\mbrc{C \mapsto \theta\mpar{x}, R \mapsto \act{set}\mpar{x}}}{\top}{\mbrc{}}} from the specification with \nmm{\OT{S}{C}{\act{set}\mpar{x}}{\mbrc{}}{\top}{\mbrc{t \gets x, c \gets 0}}} from the register and \nmm{\OT{n}{n+1}{\theta\mpar{k}}{\mbrc{}}{\top}{\mbrc{}}} from the controller.
\item \nmm{\OT{XC}{XC}{\act{tick}}{\mbrc{}}{c < t}{\mbrc{c \gets c + 1}}} which is the composition of \nmm{\OT{X}{X}{\act{tick}}{\mbrc{R \mapsto \act{tick}}}{\top}{\mbrc{}}} from the specification with \nmm{\OT{C}{C}{\act{tick}}{\mbrc{}}{c < t}{\mbrc{c \gets c + 1}}} from the register.
\item \nmm{\OT{XC}{ZS}{\act{TurnXxx}}{\mbrc{}}{c = t}{\mbrc{}}} which is the composition of \nmm{\OT{X}{Z}{\act{TurnXxx}}{\mbrc{R \mapsto \act{over}\mpar{x}, C \mapsto \delta\mpar{x}}}{\top}{\mbrc{}}} from the specification with \nmm{\OT{C}{S}{\act{over}\mpar{c}}{\mbrc{}}{c = t}{\mbrc{}}} from the register and \nmm{\OT{X}{Z}{\delta\mpar{x}}{\mbrc{}}{\top}{\mbrc{}}} from the controller.
\end{itemize}
